---
title: "MB3 Power Analysis"
author: "George & Angeline"
date: "7/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(simr)
require(tidyverse)
```

## Design


For this power analysis we will simulate 20 labs contributing 16 infants (320 participants) from 5 to 12 months of age.

Factors:

- familiarized_rule: indicates the sequence to which infants were exposed during familiarization (ABA or ABB). Infants were exposed to only one sequence, with the sequence determined by random assignment. (GK: not counterbalanced per lab?)

- test_rule (trial_type): indicates whether each test sequence followed the same rule to which the infant was familiarized or a different rule. For example, if an infant heard an ABA rule during familiarization, ABA trials would be the same trial type and ABB trials would be the different trial type. (GK: don't infants get both types of test trials?)

- trial_num: indicates the sequential order in which test trials were presented. Trial number thus ranges from 1 to 12. 

- age_mos: the infants' age in months (5.0-12.0). Age will be centered before entering the models.

- procedure: indicates the experimental method that was used to record infants’ responses to the stimuli: headturn preference procedure (HPP), central fixation (CF), or eye tracking (ET). 
- test_order: indicates which of the four pseudorandom test orders (from our provided scripts) were used to present test trials to the infant. 

- multilingual_exposure: indicates the infants' exposure to the secondary/primary language, ranging from 0% (no exposure to a secondary language) to 49% (i.e., baby hears 51% of their primary language and 49% of the secondary language). 


```{r simulate-data}
set.seed(123) # reproducible sampling

n_labs = 20
n_per_lab = 16
labID = rep(LETTERS[1:n_labs], each=n_per_lab)
subjID = 1:(n_labs*n_per_lab)

# assume each lab uses one procedure
lab_procedure = sample(c("HPP","CF","EF"), n_labs, replace=T, prob=c(.5,.3,.2))
procedure = rep(lab_procedure, each=n_per_lab)

test_order = rep(1:4, 4*n_labs) # test order is a 50/50 mix of test_rules, or not?

# randomly assigned, or counterbalanced per lab?
test_rule = sample(c("same","different"), length(subjID), replace=T)

# familiarized rule (ms says randomly assigned: we don't want counterbalanced per lab?)
familiarized_rule = sample(c("ABB","ABA"), length(subjID), replace=T)

simd <- tibble(subjID, labID, procedure, test_order, familiarized_rule, test_rule)

# uniform random vars
simd$age_mos = runif(nrow(simd), min=5.0, max=12.0)
simd$age = scale(simd$age_mos, center=T, scale=F)[,1]

# should actually be bimodal (use MB1 distro?)
simd$multilingual_exposure = runif(nrow(simd), min=0, max=.5) # 0=monolingual, .5=50% secondary language

# now just generate looking times for 12 trials per subject
for(t in 1:12) {
  simd[,paste0("trial.",t)] = rnorm(nrow(simd), mean=0, sd=1) # rexp(nrow(simd), rate=2) ?
}

siml <- simd %>% pivot_longer(cols=starts_with("trial."), 
                     names_to="trial_num", 
                     names_prefix="trial.",
                     values_to="looking_time")

# what effects do we want to simulate?
# minimally, when test_rule=="same" looking times should be longer
# maybe this effect gets stronger with age, and/or stronger with multilingual_exposure
effect_size = .3
inds_to_shift = which(siml$test_rule=="same")
siml[inds_to_shift,]$looking_time = rnorm(length(inds_to_shift), mean=-effect_size, sd=1)

```

## Model Structure

Infants' looking time (DV) ~ 1+ familiarization order (ABB vs ABA) * trial_type + age * trial_type (same rule vs different rule at test) + experimental_method (HPP vs central fixation vs eye-tracking) * trial_type + multilingual_exposure * trial_type + trial_num * trial_type + (trial_num*trial_type | subject) + (test_order | lab)

```{r model, echo=FALSE}
#m1 <- glmer(z ~ x + (1|g), family="poisson", data=simdata)
m1 <- lmer(looking_time ~ 1 + test_rule * 
             (familiarized_rule + age + procedure + multilingual_exposure + trial_num) +
             (trial_num * test_rule | subjID) + (test_order | labID), data=siml)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
