---
title: "MB3 Power Analysis"
author: "George & Angeline"
date: "7/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(simr)
require(tidyverse)
```

## Design


For this power analysis we will simulate 20 labs contributing 16 infants (320 participants) from 5 to 12 months of age.

Factors:

* *familiarized_rule*: indicates the sequence to which infants were exposed during familiarization (ABA or ABB). Infants were exposed to only one sequence, with the sequence determined by random assignment. (GK: not counterbalanced per lab?)

* *trial_type*: indicates whether each test sequence followed the same rule to which the infant was familiarized or a different rule. For example, if an infant heard an ABA rule during familiarization, ABA trials would be the *same* trial type and ABB trials would be the *different* trial type (each infant get 6 of same and 6 different)

* *trial_num*: indicates the sequential order in which test trials were presented. Trial number thus ranges from 1 to 12. 

* *age_mos*: the infants' age in months (5.0-12.0), centered in *age* column.

* *procedure*: indicates the experimental method that was used to record infants’ responses to the stimuli: headturn preference procedure (HPP), central fixation (CF), or eye tracking (ET). 
* *test_order*: indicates which of the four pseudorandom test orders (from our provided scripts) were used to present test trials to the infant. 

* *multilingual_exposure*: indicates the infants' exposure to the secondary/primary language, ranging from 0% (no exposure to a secondary language) to 49% (i.e., baby hears 51% of their primary language and 49% of the secondary language). 


```{r simulate-data}
set.seed(123) # reproducible sampling

n_labs = 20
n_per_lab = 16
labID = rep(LETTERS[1:n_labs], each=n_per_lab)
subjID = 1:(n_labs*n_per_lab)

# assume each lab uses one procedure
lab_procedure = sample(c("HPP","CF","EF"), n_labs, replace=T, prob=c(.5,.3,.2))
procedure = rep(lab_procedure, each=n_per_lab)

test_order = rep(1:4, 4*n_labs) 

# familiarized rule (ms says randomly assigned: we don't want counterbalanced per lab?)
familiarized_rule = sample(c("ABB","ABA"), length(subjID), replace=T)

simd <- tibble(subjID, labID, procedure, test_order, familiarized_rule)

# uniform random vars
simd$age_mos = runif(nrow(simd), min=5.0, max=12.0)
simd$age = scale(simd$age_mos, center=T, scale=F)[,1]

# should actually be bimodal (use MB1 distro?)
simd$multilingual_exposure = runif(nrow(simd), min=0, max=.5) # 0=monolingual, .5=50% secondary language

# now just generate looking times for 12 trials per subject
for(t in 1:12) {
  simd[,paste0("trial.",t)] = rnorm(nrow(simd), mean=0, sd=1) # rexp(nrow(simd), rate=2) ?
}

siml <- simd %>% pivot_longer(cols=starts_with("trial."), 
                     names_to="trial_num", 
                     names_prefix="trial.",
                     values_to="looking_time") 

siml$trial_num = as.numeric(siml$trial_num)

# 6 same / 6 different per child; should be according to 1 of 4 pseudorandom orders, but we're not actually modeling order effects here so just make blocks:
siml$trial_type = rep(c(rep("same", 6), rep("different", 6)), each=nrow(simd))

# what effects do we want to simulate?
# minimally, when trial_type=="same" looking times should be longer
# maybe this effect gets stronger with age, and/or stronger with multilingual_exposure
effect_size = .3
inds_to_shift = which(siml$trial_type=="same")
siml[inds_to_shift,]$looking_time = rnorm(length(inds_to_shift), mean=effect_size, sd=1)

```

## Model Structure

Infants' looking time (DV) ~ 1+ familiarization order (ABB vs ABA) * trial_type + age * trial_type (same rule vs different rule at test) + experimental_method (HPP vs central fixation vs eye-tracking) * trial_type + multilingual_exposure * trial_type + trial_num * trial_type + (trial_num*trial_type | subject) + (test_order | lab)

```{r model, echo=FALSE}
#m1 <- glmer(z ~ x + (1|g), family="poisson", data=simdata)
m1 <- lmer(looking_time ~ 1 + trial_type * 
             (familiarized_rule + age + procedure + multilingual_exposure + trial_num) +
             (trial_num * trial_type | subjID) + (test_order | labID), data=siml)
```

